---
title: "Titanic Survivor Prediction"
author: "David (Yongbock) Kwon"
output:
  html_document:
    keep_md: true
  html_notebook: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

================
Titanic Survivor Classification Prediction
=================

Importing and Manipulating Data - Feature Engineering
-----------------
```{r Import Data}
library(dplyr)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(caret)

#train and test
train <- read.csv("Datasets/train.csv", stringsAsFactors = TRUE, na.strings = "")
test <- read.csv("Datasets/test.csv", stringsAsFactors = TRUE, na.strings = "")


#creating survived variables in test set and combinining train and test
test$Survived <- NA
dat <- rbind(train,test)

```

Survived and Pclass
--------------------
```{r Survived and Pclass}
#convert survived and pclass to factor variable
dat$Survived <- as.factor(dat$Survived)
dat$Pclass <- as.factor(dat$Pclass)
#Survived : 1 / no Survived : 0

#Bar graph for Pclass vs Survived
dat %>% filter(!is.na(Survived)) %>%
  ggplot(aes(x=Pclass, fill=Survived))+
  geom_bar()+
  geom_label(stat="count", 
             position=position_stack(0.5), 
             aes(label=..count..))+
  ggtitle("Pclass vs Survived")

#Bar graph for Sex vs Survived
dat %>% filter(!is.na(Survived)) %>%
  ggplot(aes(x=Sex, fill=Survived))+
  geom_bar()+
  geom_label(stat="count", 
             position=position_stack(0.5), 
             aes(label=..count..))+
  ggtitle("Sex vs Survived")


dat %>% filter(!is.na(Survived)) %>%
  ggplot(aes(x=Pclass, fill=Survived))+
  geom_bar()+
  geom_label(stat="count", 
             position=position_stack(0.5), 
             aes(label=..count..))+
  ggtitle("Pclass vs Survived by Sex")+
  facet_grid(~Sex)

#In Pclass 1 and 2, obviously male mostly not survived and female survived
#In Pclass 3, male mostly not survived, but female hard to predict whether surv or not


```

From Cabin, Cabin.ox
--------------------
```{r Cabin.ox}
#Cabin NA values -> 0, otherwise 1
dat$Cabin.ox <- as.factor(ifelse(is.na(dat$Cabin), 0, 1))
table(dat$Cabin.ox)
#no cabin : 0 / cabin : 1

dat %>% filter(!is.na(Survived)) %>%
  ggplot(aes(x=Cabin.ox, fill=Survived))+
  geom_bar()+
  geom_label(stat="count", 
             position=position_stack(0.5), 
             aes(label=..count..))+
  ggtitle("Cabin.ox vs Survived")

dat %>% filter(!is.na(Survived)) %>%
  ggplot(aes(x=Cabin.ox, fill=Survived))+
  geom_bar()+
  geom_label(stat="count", 
             position=position_stack(0.5), 
             aes(label=..count..))+
  ggtitle("Cabin.ox vs Survived by Sex")+
  facet_grid(~Sex)

#If Cabin is not NA, then more likely survived
#no cabin likely not survived

dat %>% filter(!is.na(Survived)) %>%
  ggplot(aes(x=Cabin.ox, fill=Survived))+
  geom_bar()+
  geom_label(stat="count", 
             position=position_stack(0.5), 
             aes(label=..count..))+
  ggtitle("Cabin.ox vs Survived by Pclass")+
  facet_grid(~Pclass)

#Also, notice Pclass 1 people mostly have cabin
#Pclass 2 and 3 not

dat %>% filter(!is.na(Survived)) %>%
  ggplot(aes(x=Fare, fill=Cabin.ox))+
  geom_density(alpha=0.3)+
  ggtitle("Fare vs Cabin.ox")

dat %>% filter(!is.na(Survived)) %>%
  ggplot(aes(x=Cabin.ox, y=Fare))+
  geom_boxplot()+
  ggtitle("Fare vs Cabin.ox")

#Fare difference by Cabin.ox

```

Function to make prop.table
--------------------------
```{r prop.func}
#creating function to make prop.table

prop.func <- function(predictor){
  prop.tab <- data.frame(
    prop.table(
      matrix(
        c(table(dat[1:891,predictor], dat$Survived[1:891])[,1],
          table(dat[1:891,predictor], dat$Survived[1:891])[,2]),
        ncol=2),
      1))
  colnames(prop.tab) <- c("no surv", "surv")
  rownames(prop.tab) <- c(levels(dat[,predictor]))
  
  return(prop.tab)
}
```


From Cabin, deck.surv
-----------------
```{r deck.surv}
#deck from Cabin
dat$deck <- as.factor(ifelse(is.na(substr(dat$Cabin,1,1)), "no", substr(dat$Cabin,1,1)))

which(dat$deck == "T") #the element where is in traing set.. lets replace this to something else

dat %>% 
  subset(select = -c(PassengerId)) %>% 
  filter(!is.na(Survived)) %>% 
  group_by(deck) %>% 
  summarise(count = n(),
            mean = mean(Fare))
#mean of Fare for deck "T" is close to the mean of Fare for deck "A" 
#replace "T" to "A"

dat$deck[dat$deck=="T"] <- "A"
dat$deck <- as.factor(as.character(dat$deck))

summary(dat$deck)

#proportional bar graph 
dat %>% filter(!is.na(Survived)) %>%
  ggplot(aes(x=deck, fill=Survived))+
  geom_bar(position = "fill")+
  ggtitle("Deck Survival rate by Proportion")

#count bar graph without no deck
dat %>% filter(!is.na(Survived) & deck != "no") %>% 
  ggplot(aes(x=deck, fill=Survived)) + 
  geom_bar() + 
  geom_label(stat = "count", 
             position = position_stack(0.5), 
             aes(label= ..count..))+
  ggtitle("Deck Survival by count without no deck")

table(dat$deck[1:891], dat$Survived[1:891])

deck.prop <- prop.func("deck")

#proportional deck table
deck.prop

#we might want to group up B/D/E together (which have high prob for survived)
#so, B/C/D/E/F -> high prob surv rate deck
#    A/G/no -> low prob surv rate
dat$deck <- as.character(dat$deck)

dat$deck.surv <- NA
for(i in 1:nrow(dat)){
  if(dat$deck[i] %in% c("B", "C", "D", "E","F")){
    dat$deck.surv[i] <- "high"
  }
  if(dat$deck[i] %in% c("no", "A", "G")){
    dat$deck.surv[i] <- "low"
  }
}

table(dat$deck.surv)

dat$deck.surv <- as.factor(dat$deck.surv)

dat <- dat %>% subset(select=-c(deck))
```



From Cabin, cabin.freq.surv
--------------------------
```{r cabin.freq.surv}
#cabin frequency.. might have relationship between cabin freq
cabin.freq <- data.frame(table(dat$Cabin))


dat$cabin.freq <- NA
for(i in 1:nrow(dat)){
  if(dat$Cabin[i] %in% cabin.freq$Var1){
    dat$cabin.freq[i] <- cabin.freq$Freq[cabin.freq$Var1==dat$Cabin[i]]
  }
  else{
    dat$cabin.freq[i] <- 0
  }
}
dat$cabin.freq <- as.factor(dat$cabin.freq)
summary(dat$cabin.freq)


#proportional bar graph
dat %>% filter(!is.na(Survived)) %>% 
  ggplot(aes(x=cabin.freq, fill=Survived)) + 
  geom_bar(position = "fill")+
  ggtitle("Cabin Frequency Survival Rate by Proportion")

#bar graph without no cabin
dat %>% filter(!is.na(Survived) & cabin.freq != 0) %>% 
  ggplot(aes(x=cabin.freq, fill=Survived)) + 
  geom_bar() + 
  geom_label(stat = "count", 
             position = position_stack(0.5), 
             aes(label= ..count..))+
  ggtitle("Cabin Frequency Survival by Count")


table(dat$cabin.freq[1:891], dat$Survived[1:891])

cabin.freq.prop <- prop.func("cabin.freq")


cabin.freq.prop

#no cabin barely survived
#cabin freq 1 / 2 / 3 / 4 / 5 more likely surv

#no cabin , cabin freq 6 -> low
#cabin freq 1,2,3,4,5    -> high

dat$cabin.freq.surv <- NA

for(i in 1:nrow(dat)){
  if(dat$cabin.freq[i] %in% c(1,2,3,4,5)){
    dat$cabin.freq.surv[i] <- "high"
  }
  if(dat$cabin.freq[i] %in% c(0,6)){
    dat$cabin.freq.surv[i] <- "low"
  }
}

dat$cabin.freq.surv <- as.factor(dat$cabin.freq.surv)
table(dat$cabin.freq.surv)

dat <- subset(dat, select = -c(Cabin, cabin.freq))
```


Dealing with NA values in Embarked and Fare
--------------------------------
```{r NA value in Embarked / Fare}
#Gender -> male = 0, female = 1
dat$Sex <- as.factor(ifelse(dat$Sex == "male", 0, 1))

dat[is.na(dat$Embarked),]

#Pclass = 1 / Sex = Female / have cabin / 
#deck surv rate high / cabin freq surv rate high
dat %>% 
  filter(Pclass == 1 & 
           Sex == 1 & 
           Cabin.ox==1 & 
           deck.surv == "high" & 
           cabin.freq.surv == "high" & 
           SibSp == 0 & 
           Parch == 0) %>% group_by(Embarked) %>%
  summarise(count = n(),
            mean = mean(Fare),
            min = min(Fare),
            max = max(Fare))



#Na value for Embarked
dat$Embarked[is.na(dat$Embarked)] <- "C"

dat[is.na(dat$Fare),]

summary(aov(Fare~Cabin.ox, dat))
summary(aov(Fare~Pclass, dat))

#NA value for Fare
dat[dat$Pclass == 3,] %>% 
  group_by(Embarked, Cabin.ox, Pclass) %>% 
  summarise(mean = mean(Fare, na.rm=TRUE))

#Pclass 3 / Embarked S / no cabin
#mean of Pclass 3 and Embarked S, and no cabin is 14.5

dat$Fare[is.na(dat$Fare)] <- 14.5
```


From Ticket, ticket.alone
--------------------------
```{r ticket.alone}
#Ticket
ticket.alone <- data.frame(table(dat$Ticket))

dat$ticket.alone <- NA
for(i in 1:nrow(dat)){
  if(dat$Ticket[i] %in% ticket.alone$Var1[ticket.alone$Freq==1]){
    dat$ticket.alone[i] <- 0
  }
  if(dat$Ticket[i] %in% ticket.alone$Var1[ticket.alone$Freq>1]){
    dat$ticket.alone[i] <- 1
  }
}

table(dat$ticket.alone)

dat$ticket.alone <- as.factor(dat$ticket.alone)
```


From Ticket, ticket.let.surv
-------------------------
```{r ticket.let.surv}
#ticket by first letter
dat$ticket.let <- substr(dat$Ticket, 1,1)

ticket.let <- data.frame(table(dat$ticket.let))


#proportional bar graph
dat %>% filter(!is.na(Survived) ) %>% 
  ggplot(aes(x=ticket.let, fill=Survived)) + 
  geom_bar(position = "fill")+
  ggtitle("Ticket by first letter surv rate by proportional")

dat %>% filter(!is.na(Survived) ) %>% 
  ggplot(aes(x=ticket.let, fill=Survived)) + 
  geom_bar() + 
  geom_label(stat = "count", 
             position = position_stack(0.5), 
             aes(label= ..count..))+
  ggtitle("Ticket by first letter surv rate by count")


table(dat$ticket.let[1:891], dat$Survived[1:891])

dat$ticket.let <- as.factor(dat$ticket.let)

ticket.let.prop <- prop.func("ticket.let")
ticket.let.prop


dat$ticket.let <- as.factor(dat$ticket.let)

die <- rownames(ticket.let.prop[ticket.let.prop$`no surv`>=0.5,])
surv <- rownames(ticket.let.prop[ticket.let.prop$`no surv`<0.5,])

dat$ticket.let <- as.character(dat$ticket.let)
dat$ticket.let.surv <- NA
for(i in 1:nrow(dat)){
  if(dat$ticket.let[i] %in% die){
    dat$ticket.let.surv[i] <- "low"
  }
  if(dat$ticket.let[i] %in% surv){
    dat$ticket.let.surv[i] <- "high"
  }
}

dat$ticket.let.surv <- as.factor(dat$ticket.let.surv)
summary(dat$ticket.let.surv)


dat <- dat %>% subset(select =-c(Ticket, ticket.let))
```


Creating family variable
-----------------------
```{r family}
#family size (if family = 1, then it's alone)
dat$family <- dat$SibSp + dat$Parch + 1
#1 == alone

dat <- subset(dat, select = -c(SibSp, Parch))
```


From Name, name and surname.freq.surv
Dealing with NA values in Age
---------------------------
```{r name / surname.freq.surv / NA in Age}
#converting names
dat <- dat %>% 
  mutate(name = sub("\\..*$", "", sub("^.*, ", "", Name)), 
         surname = sub(",.*$", "", Name))

summary(as.factor(dat$name))
summary(as.factor(dat$surname))
#name first

dat %>% 
  group_by(name, Sex) %>% 
  summarise(mean = mean(Age, na.rm=TRUE),
            min = min(Age, na.rm=TRUE),
            max = max(Age, na.rm=TRUE),
            count = n())


#Master / Miss / Mr / Mrs
#Matser seems obvious young male
#Mr teenage to old male
#Miss and Mrs female in range young to old

#Age first.. to predict name by age

dat %>% filter(is.na(Age)) %>% group_by(name,Sex) %>% tally()

#dealing with Dr
dat %>% filter(name == "Dr")

dat$Age[which(dat$name == "Dr" & is.na(dat$Age))] <- mean(dat$Age[which(dat$name == "Dr")], na.rm=TRUE)

#dealing with Ms
dat %>% filter(name == "Ms")

dat$Age[which(dat$name == "Ms" & is.na(dat$Age))] <- mean(dat$Age[which(dat$name == "Ms")], na.rm=TRUE)

dat$name <- as.character(dat$name)
dat$surname <- as.character(dat$surname)


summary(aov(Age~Pclass, dat))
summary(aov(Age~name, dat))
#I use Pclass and name to predict NA values in Age
#replacing NA's of Age with the mean by name and Pclass, respectively

dat %>% filter(is.na(Age)) %>% group_by(name,Pclass) %>% tally()

dat[dat$name %in% c("Mr", "Miss", "Mrs", "Master"),] %>% 
  group_by(name, Pclass) %>% 
  summarise(count = n(), 
            mean = mean(Age, na.rm=TRUE),
            min = min(Age, na.rm=TRUE),
            max = max(Age, na.rm=TRUE))

for(i in 1:nrow(dat)){
  if(is.na(dat$Age[i])){
    #Master
    if(dat$name[i] == "Master" & dat$Pclass[i] == 3){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Master" & dat$Pclass == 3)], na.rm=TRUE)
    }
    
    #Miss
    if(dat$name[i] == "Miss" & dat$Pclass[i] == 1){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Miss" & dat$Pclass == 1)], na.rm=TRUE)
    }
    if(dat$name[i] == "Miss" & dat$Pclass[i] == 2){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Miss" & dat$Pclass == 2)], na.rm=TRUE)
    }
    if(dat$name[i] == "Miss" & dat$Pclass[i] == 3){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Miss" & dat$Pclass == 3)], na.rm=TRUE)
    }
    
    #Mr
    if(dat$name[i] == "Mr" & dat$Pclass[i] == 1){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Mr" & dat$Pclass == 1)], na.rm=TRUE)
    }
    if(dat$name[i] == "Mr" & dat$Pclass[i] == 2){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Mr" & dat$Pclass == 2)], na.rm=TRUE)
    }
    if(dat$name[i] == "Mr" & dat$Pclass[i] == 3){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Mr" & dat$Pclass == 3)], na.rm=TRUE)
    }
    
    #Mrs
    if(dat$name[i] == "Mrs" & dat$Pclass[i] == 1){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Mrs" & dat$Pclass == 1)], na.rm=TRUE)
    }
    if(dat$name[i] == "Mrs" & dat$Pclass[i] == 2){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Mrs" & dat$Pclass == 2)], na.rm=TRUE)
    }
    if(dat$name[i] == "Mrs" & dat$Pclass[i] == 3){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Mrs" & dat$Pclass == 3)], na.rm=TRUE)
    }
    
    #Ms
    if(dat$name[i] == "Ms" & dat$Pclass[i] == 3){
      dat$Age[i] <- mean(dat$Age[which(dat$name == "Ms" & dat$Pclass == 3)], na.rm=TRUE)
    }
  }
}


#dealing with other names
dat$name[!dat$name %in% c("Mr", "Miss", "Mrs", "Master") ]


dat %>% filter(!name %in% c("Mr", "Miss", "Mrs", "Master")) %>% 
  group_by(name, Sex) %>% 
  summarise(count = n(), 
            mean = mean(Age),
            min = min(Age, na.rm=TRUE),
            max = max(Age, na.rm=TRUE))

dat[dat$name %in% c("Mr", "Miss", "Mrs", "Master"),] %>% 
  group_by(name) %>% 
  summarise(count = n(), 
            mean = mean(Age, na.rm=TRUE),
            min = min(Age, na.rm=TRUE),
            max = max(Age, na.rm=TRUE))

#Master max age 14.5
#Master -> young male : sex==male & Age < 14.5
#Mr -> adult male :     sex==male & Age > 14.5
#Miss -> adult female : sex==female & Age < 14
#Mrs -> adult female :  sex==female & Age > 14

for(i in 1:nrow(dat)){
  if(!is.na(dat$Age[i])){
    if(!dat$name[i] %in% c("Mr", "Miss", "Mrs", "Master")){
      if(dat$Sex[i] == 0 & dat$Age[i] <= 14.5){
      dat$name[i] = "Master"
      }
      if(dat$Sex[i] == 0 & dat$Age[i] > 14.5){
      dat$name[i] <- "Mr"
      }
      if(dat$Sex[i] == 1 & dat$Age[i] < 14){
      dat$name[i] <- "Miss"
      }
      if(dat$Sex[i] == 1 & dat$Age[i] > 14){
      dat$name[i] <- "Mrs"
      }
    }
  }
}


dat$name <- as.factor(as.character(dat$name))

table(dat$name)

#surname frequency
surname.freq <- data.frame(table(dat$surname))

dat$surname.freq <-NA

for(i in 1:nrow(dat)){
  for(j in 1:11){
    if(dat$surname[i] %in% surname.freq$Var1[surname.freq$Freq == j]){
      dat$surname.freq[i] <- j
    }
  }
}

dat$surname.freq <- as.factor(dat$surname.freq)


#bar graph
dat %>% filter(!is.na(Survived)) %>% 
  ggplot(aes(x=surname.freq, fill=Survived)) + 
  geom_bar(position = "fill")+
  ggtitle("Surname Frequency Survival Rate by Proportion")

dat %>% filter(!is.na(Survived)) %>% 
  ggplot(aes(x=surname.freq, fill=Survived)) + 
  geom_bar() + 
  geom_label(stat = "count", position = position_stack(0.5), aes(label= ..count..))+
  ggtitle("Surname Frequency Survival by Count")


table(dat$surname.freq[1:891], dat$Survived[1:891])



surname.freq.prop <- prop.func("surname.freq")

surname.freq.prop
#notice that surname.freq 2,3 is likely hard to predict 
#however, more the surname.freq increased from 4 to 11, they are more likely not survived

#therefore, low surv rate -> 1,4,5,6,7,8,11
#unknown -> 2,3

dat$surname.freq <- as.character(dat$surname.freq)

dat$surname.freq.surv <- NA
for(i in 1:nrow(dat)){
  if(dat$surname.freq[i] %in% c(1,4,5,6,7,8,11)){
    dat$surname.freq.surv[i] <- "low"
  }
  if(dat$surname.freq[i] %in% c(2,3)){
    dat$surname.freq.surv[i] <- "unknown"
  }
}
dat$surname.freq.surv <- as.factor(dat$surname.freq.surv)

table(dat$surname.freq.surv)

dat <- subset(dat, select=-c(surname.freq, Name, surname))

summary(dat)
```



Investigating correlation or relationship between each variables in our dataset
-----------------------------
```{r correlation or relationship between each vars}
#Let's see the correlation or relationship between each variables in our dataset

#factor vs factor - chisq test : null H0 = two factor variables are independent
#factor vs numeric - anova test : null H0 = at least one factor has different mean than others
#numeric vs numeric - correlation : linear relationship between vars, 
#more than 0.5 means they have some relationship to each other

relationship.test <- function(variables, dummy.data, data){

  for(i in variables){
    for(j in variables){
      
      #factor vs factor : chisq.test
      if(is.factor(data[,i])){
        if(is.factor(data[,j])){
          dummy.data[dummy.data$cols == i,j] <- round(chisq.test(data[,i], data[,j])$p.value,3)
          }
      }
      
      #factor vs numeric : anova
      if(is.factor(data[,i])){
        if(is.numeric(data[,j])){
          dummy.data[dummy.data$cols == i,j] <-
            round(summary(aov(data[,j]~data[,i]))[[1]][["Pr(>F)"]][[1]],3)
          }
      }
      if(is.numeric(data[,i])){
        if(is.factor(data[,j])){
          dummy.data[dummy.data$cols == i,j] <- 
            round(summary(aov(data[,i]~data[,j]))[[1]][["Pr(>F)"]][[1]],3)
          }
        }
    
      #numeric vs numeric : correlation
      if(is.numeric(data[,i])){
        if(is.numeric(data[,j])){
          dummy.data[dummy.data$cols == i,j] <- round(cor(data[,i], data[,j]),3)
          }
        }
    }
  }
  
  return(dummy.data)
}

#creating variables
variables <- colnames(dat)[2:ncol(dat)]

#dummy data
test.data <- data.frame(cols = variables)

data.pval <- relationship.test(variables, test.data, dat)
data.pval


#factor vs factor : if <0.05 (p value), highly dependent, if not, independent
#factor vs numeric : if <0.05, at least one factor has different mean than others. 
#if not, all factor has similar mean (non linear)
#numeric vs numeric : if <0.5, low correlation, if not, high correlation
```


Creating familyGroup from investigation of relationship between each variables
----------------------------
```{r familyGroup}
#Lets make family to be better predictor

tr <- rpart(Survived~family, dat)
tr
rpart.plot(tr)


dat %>% filter(!is.na(Survived)) %>% 
  ggplot(aes(x=family, fill=Survived))+
  geom_density(alpha = 0.3)+
  geom_vline(xintercept=c(1.5, 4.5))
#1.5 and 4.5

dat$familyGroup <- as.factor(ifelse(dat$family < 1.5, "alone",
                                    ifelse(dat$family > 1.5 & dat$family < 4.5, "small fam", "large fam")))

table(dat$familyGroup)


variables <- colnames(dat)[2:ncol(dat)]
test.data <- data.frame(cols = variables)
test.data

data.pval <- relationship.test(variables, test.data, dat)
data.pval[,1:2]


dat <- dat %>% subset(select=-c(PassengerId, family))

summary(dat)
```

Splitting train and test set to start modeling
-----------------------------------------
```{r splitting train and test}

#train / test
training <- dat %>% filter(!is.na(Survived))
testing <- dat %>% filter(is.na(Survived))

summary(training)
summary(testing)

#we have 14 predictors. 
#we might want to remove some predictors that have low importance while modeling
```

From Cabin..
  - Cabin.ox : Cabin NA = 0 or Cabin = 1
  - deck.surv : extract the first letter of cabin, with the probability of survival for the deck, splitted into 2 groups, which are high / low
  - cabin.freq.surv : 2 groups by surv rate with cabin frequency

from Ticket..
  - ticket.alone : unique ticket = 0 other 1
  - ticket.let.surv : with the first letter of ticket, splitted into 2 groups by surv rate of the ticket letter

from Name..
  - name : Master / Miss / Mr / Mrs
  - surname.freq.surv : groups by surv rate with surname frequency



Caret - Cross Validation
Creating useful function for modeling
---------------------
```{r modeling / accuracy / trainControl function}
#creating function for Caret modeling

model <- function(method, training, control,grid,...){

  if(is.null(grid)){
    model.fit <- train(Survived~.,
                     data = training,
                     method = method,
                     trControl = control,
                     ...)
    return(model.fit)
  }

  else{
    model.fit <- train(Survived~.,
                     data = training,
                     method = method,
                     trControl = control,
                     tuneGrid = grid,
                     ...)
    return(model.fit)
  }
}

#accuracy of model
acc <- function(pred, act, data){
  return(sum(diag(table(pred, act)))/nrow(data))
}

#10 folds cv
control <- trainControl(method = "cv", number = 10)

```


I will use Random Forest / Gradient Boosting Method / Support Vector Machine with kernel radial


Random Forest
-------------
```{r Random Forest}

#typical mtry in classification = sqrt(# of predictors)
rf.fit <- train(Survived~., data = training,
                method="rf", trControl = control, 
                ntree=500, importance = TRUE, 
                tuneGrid = expand.grid(mtry = round(sqrt(ncol(training)-1))))

rf.fit

plot(rf.fit$finalModel)
varImp(rf.fit)

rf.fit.result <- data.frame(rf.fit$finalModel$importance[,"MeanDecreaseAccuracy"])
colnames(rf.fit.result) <- "MeanDecreaseAccuracy"

rf.fit.result

rf.fit.result %>% mutate(name = rownames(rf.fit.result)) %>% 
  arrange(MeanDecreaseAccuracy) %>%
  ggplot(aes(x=reorder(name, MeanDecreaseAccuracy),y=MeanDecreaseAccuracy))+
  geom_point()+
  theme(axis.text.x = element_text(angle=90))
#remove Embarked / surname.freq.surv 


#tuning parameter mtry and ntree by cross validation
#typical mtry is sqrt(# of predictor)
#ntree: in small dataset -> 100 in large dataset -> 500~1000 sufficient
#larger ntree is more stable, but takes long time
rf.grid <- expand.grid(mtry = seq(2,10, by=2))

rf.acc <- data.frame(ntree = seq(100,1000, by=100), minacc = NA, acc = NA)

for(i in seq(100, 1000, by=100)){
  rf.fit <- train(Survived~., data=training %>% subset(select = -c(Embarked, surname.freq.surv)),
                  method = "rf", trControl = control, 
                  ntree=i, tuneGrid = rf.grid, importance = TRUE)
  rf.acc[rf.acc$ntree == i,2] <- max(rf.fit$results$Accuracy) -
    rf.fit$results$AccuracySD[which.max(rf.fit$results$Accuracy)]
  rf.acc[rf.acc$ntree == i,3] <- max(rf.fit$results$Accuracy)
}

rf.acc

ggplot(rf.acc, aes(x=ntree, y=acc))+
  geom_line()+
  geom_point()

g.ntree <- rf.acc$ntree[which.max(rf.acc$minacc)]
g.ntree
#I will choose the ntree that has maximum value of minacc = max accuracy - accuracy sd

rf.model <- train(Survived~., 
                  data=training %>% subset(select=-c(Embarked, surname.freq.surv)), 
                  method = "rf", trControl = control, 
                  ntree=g.ntree, tuneGrid = rf.grid, importance=TRUE)

rf.model

plot(rf.model$finalModel)
max(rf.model$results$Accuracy)
#about 83%

varImp(rf.model)


rf.model.result <- data.frame(rf.model$finalModel$importance[,"MeanDecreaseAccuracy"])
colnames(rf.model.result) <- "MeanDecreaseAccuracy"

rf.model.result

rf.model.result %>% mutate(name = rownames(rf.model.result)) %>% 
  arrange(MeanDecreaseAccuracy) %>%
  ggplot(aes(x=reorder(name, MeanDecreaseAccuracy),y=MeanDecreaseAccuracy))+
  geom_point()+
  theme(axis.text.x = element_text(angle=90))


rf.minacc <- max(rf.model$results$Accuracy) - 
  rf.model$results$AccuracySD[which.max(rf.model$results$Accuracy)] 
rf.minacc
#about 80%

#predict on real test
rf.pred <- predict(rf.model, training)

confusionMatrix(rf.pred, training$Survived)
#93.15%

#training accuracy - cv accuracy
acc(rf.pred, training$Survived, training) - max(rf.model$results$Accuracy)
#0.0987


```




Gradient Boosting Method
-------------

```{r GBM}
#modeling without tuning parameter
boost.model <- train(Survived~.,
                   data = training,
                   method = "gbm",
                   verbose = FALSE,
                   trControl = control,
                   tuneGrid = NULL)

boost.model
summary(boost.model$finalModel)
#surname.freq.surv / Embarked



#Grid Search
#I put relatively large value of shrinkage to prevent overfitting
boost.grid <- expand.grid(n.trees = seq(100,6000, by=150),
                          interaction.depth = c(1,2,3,4),
                          shrinkage = c(0.01,0.1),
                          n.minobsinnode = c(10))

#modeling
boost.model <- train(Survived~.,
                   data = training %>% 
                     subset(select = -c(Embarked, surname.freq.surv)),
                   method = "gbm",
                   verbose = FALSE,
                   trControl = control,
                   tuneGrid = boost.grid)

boost.model$bestTune

max(boost.model$results$Accuracy) 
#84.44%

boost.minacc <- max(boost.model$results$Accuracy) - 
  boost.model$results$AccuracySD[which.max(boost.model$results$Accuracy)] 
boost.minacc
#81.28%

plot(boost.model)

summary(boost.model$finalModel)

boost.model$finalModel$tuneValue$n.trees

#predict on training
boost.pred <- predict(boost.model, training, 
                      n.trees=boost.model$finalModel$tuneValue$n.trees)


confusionMatrix(boost.pred, training$Survived)
#88.78%

acc(boost.pred, training$Survived, training) - max(boost.model$results$Accuracy)
#0.0437



```



SVM - kernel radial
-------------------

```{r SVM}

svm.radial <- model("svmRadial", training, control, grid = NULL, tuneLength = 10)
svm.radial 

max(svm.radial$results$Accuracy)
#83.16%

varImp(svm.radial)
#name and Age

#Grid Search for tuning parameter
svm.grid <- expand.grid(sigma = seq(0.01,0.1, by=0.01),
                        C = seq(0.01,2.01,by=0.25))

svm.radial <- model("svmRadial", training %>% subset(select = -c(name, Age)), 
                    control, 
                    grid = svm.grid)

svm.radial$bestTune
max(svm.radial$results$Accuracy)
#0.8160

#on training
svm.radial.pred <- predict(svm.radial, training)

confusionMatrix(svm.radial.pred, training$Survived)
#0.8395

acc(svm.radial.pred, training$Survived, training) - max(svm.radial$results$Accuracy)
#0.0235
```






Ensembling models in a dataset
-------------
```{r Ensembling models}


#prediction on test

rf.test.pred <- predict(rf.model, testing)
boost.test.pred <- predict(boost.model, testing)
svm.radial.pred <- predict(svm.radial, testing)

ensembled.test <- data.frame(PassengerId = test$PassengerId, 
                             rf = rf.test.pred, 
                             boost= boost.test.pred, 
                             svm = svm.radial.pred)

#Take average of the predicting value by 3 models : Random Forest / Gradient Boosting / SVM - Radial
ensembled.test$mean <- as.factor(round((as.numeric(ensembled.test$rf) + 
                                          as.numeric(ensembled.test$boost) + 
                                          as.numeric(ensembled.test$svm) - 3)/3))

ensembled.test$PassengerId <- as.character(ensembled.test$PassengerId)

summary(ensembled.test)

```



Creating submission
--------------
```{r Submission}

final.pred <- ensembled.test$mean
final.pred

final <- data.frame(PassengerId = test$PassengerId, Survived = final.pred)

head(final)

#write.csv(final, "/Users/DavidKwon/Desktop/Practice/Kaggle/Titanic/final.csv", row.names = FALSE)

```

Public Score - The public score is different by seed, but it's about 78~79%


